---
title: "PA1_templete"
author: "Zuraida"
date: "Monday, October 19, 2015"
output: html_document
---


#Loading and preprocessing the data

Here is a code chunk for loading and preprocessing the data.
```{r load and preprocess}
getwd()
setwd("D:/R assignment/Reproducible research")
##setwd("D:/Bid data cousera/Reproducible research")
MyData <- read.csv(file="activity.csv", header=TRUE, sep=",")
```

#Mean total number of steps taken per day

Here is the code chunk for sum, mean and median total number of steps taken per day.
```{r mean total number of steps taken per day}


sum_daily <- aggregate(MyData$steps, list(date=MyData$date), sum, na.rm = TRUE)
mean_daily <- aggregate (MyData$steps, list(date=MyData$date), mean, na.rm = TRUE)
mean_daily
median_daily <- aggregate (MyData$steps, list (date=MyData$date), median,na.rm = TRUE)
median_daily
```

Here is the table for means
```{r show means, results="asis"}
library(datasets)
library(xtable)

xt <- xtable(mean_daily)
print(xt, type="html")
```

Here is the table for medians
```{r show median, results="asis"}
library(datasets)
library(xtable)

xt <- xtable(median_daily)
print(xt, type="html")
```

#Average daily activity pattern

Here is the code chunck for average daily pattern.
```{r average daily pattern}
average_steps_interval <- aggregate(MyData$steps, list(interval=MyData$interval), mean, na.rm = TRUE)

plot(average_steps_interval$interval, average_steps_interval$x, xlab= "5 Minutes Interval", ylab= "average steps across all days", type='l', col='red') 
Max <- max(average_steps_interval$x)
Max_interval <- average_steps_interval[which(average_steps_interval$x == Max),1]
Max_interval
```

The 5-minute interval, on average across all the days in the dataset which contains the maximum number of steps is `r Max_interval`.

#Imputing missing values

Here is the code chunk to manage missing values in the dataset. 

```{r handle missing values}
sum_missing <- sum(is.na(MyData))

m <- dim(MyData)
n <- m[1]

##create new dataset 
NewData <- MyData
NewData[is.na(NewData)] <- 9999


for(i in 1:n){
  if (NewData[i,1]== 9999)
  {
    j <- NewData[i,3]
    
    Mean <-average_steps_interval[which(average_steps_interval$interval == j),2]
    NewData[i,1] <- Mean
  }
}

##Total number of steps taken each day
TotalStepDaily <- aggregate(NewData$steps, list(date=NewData$date), sum)
hist(TotalStepDaily$x)


##Mean and median total number of steps taken per day
MeanDailyNew <- aggregate (NewData$steps, list(date=NewData$date), mean)
MedianDailyNew <- aggregate (NewData$steps, list (date=NewData$date), median)

```

The total number of missing values in the original dataset is `r sum_missing` 

The mean of the total number of steps taken per day is as follows.

```{r show mean new, results="asis"}
library(datasets)
library(xtable)

xt <- xtable(MeanDailyNew)
print(xt, type="html")
```

The median of the total number of steps taken per day is as follows.

```{r show median new, results="asis"}
library(datasets)
library(xtable)

xt <- xtable(MedianDailyNew)
print(xt, type="html")
```

It can be seen that there is difference between these values and the estimates from the first part of the assignment. 
The impact from filling in the missing values is we obtain full estimates for all the days in which the previous estimates do not have. 


#The differences in activity patterns between weekdays and weekends

Here is the code chunck for the evaluating the differences between weekend and weekday. 

```{r differences weekend and weekday}
#install.packages('timeDate')
#library(timeDate)
NewData$date <- as.Date(NewData$date)
weekdays1 <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
NewData$daytype <- c('weekend', 'weekday')[(weekdays(NewData$date) %in% weekdays1)+1L]

MeanDayTypeNew <- aggregate (NewData$steps, list(day=NewData$daytype, interval=NewData$interval), mean)

library('ggplot2')
averages <- aggregate(steps ~ interval + daytype, data=NewData, mean)
ggplot(averages, aes(interval, steps)) + geom_line() + facet_grid(daytype ~ .) +
    xlab("5-minute interval") + ylab("Number of steps")
```
